---
title: "CodingNotes"
output: html_document
---

```{r}
library(ggplot2)
library(drc)
library(tidyverse)
```

manually converting fahrenheit to celsius

```{r}
(5*(32 - 32)/9)
(5*(80 - 32)/9)
```

converting it into a function

```{r}
F_to_C <- function(fahrenheit_temp){
  celsius <- (5*(fahrenheit_temp - 32)/9)
  return(celsius)
}

F_to_C(32)
F_to_C(80)
```

function to convert celsius to fahrenheit

```{r}
C_to_F <- function(celsius_temp){
  fahrenheit <- (celsius_temp*(9/5)+32)
  return(fahrenheit)
}

C_to_F(0)
C_to_F(26.66667)
```

Iterations functions in base R

the rep() function (allows you to repeat elements easily)
```{r}
rep("A", 3) # repeats A three times 
```

```{r}
rep(c("A", "B"), 5) # repeats A and B, 5 times 
```

```{r}
rep(c(1,2,3,4), times = 4) # repeats 1,2,3,4, 4 times 
```

```{r}
rep(c(1,2,5,2), times = 4, each = 4) # repeats 1 four times, 2 four times, 5 four times, and 2 four times. 
```


seq() command allows you to write sequences of numbers easily
```{r}
seq(from = 1, to = 7) # sequence of numbers 1 to 7
```

```{r}
seq(from = 0, to = 10, by = 2) # sequence of numbers from 0 to 10 by 2s
```


combined seq() and rep()
```{r}
rep(seq(from = 0, to = 10, by = 2), times = 3, each = 2)
```


seq_along() function (allows you to generate a sequence of numbers based on non-integer values.)
```{r}
LETTERS
seq_along(LETTERS)
```


The for loop

```{r}
for (i in 1:10) {
print(i*2)  
}
```

```{r}
for (i in -30:10){
  result <- F_to_C(i)
  print(result)
}
```


```{r}
celcius.df <- NULL 
for (i in -30:100){   
result_i <- data.frame(F_to_C(i), i)   
celcius.df <- rbind.data.frame(celcius.df, result_i) 
}
```


practical example

```{r}
# read in the data
EC50.data <- read.csv("EC50_all.csv")
```

calculating manually
```{r}
isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~ 
        EC50.data$conc[EC50.data$is == "ILSO_5-41c"], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    summary.fit <- data.frame(summary(isolate1)[[3]])
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    
EC50
```

The loop
```{r}
nm <- unique(EC50.data$is)
```


```{r}
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    summary.fit <- data.frame(summary(isolate1)[[3]])
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    EC50
}
```

saving the results in a new dataframe

```{r}
EC50.ll4 <- NULL 
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    summary.fit <- data.frame(summary(isolate1)[[3]])
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    EC50
    isolate.ec_i <- data.frame(nm[[i]], EC50)
    colnames(isolate.ec_i) <- c("Isolate", "EC50") 
    EC50.ll4 <- rbind.data.frame(EC50.ll4, isolate.ec_i)
}
```

plotting final dataset

```{r}
ggplot(EC50.ll4, aes(x = EC50)) + 
  geom_histogram() + 
  theme_classic()
```

Using map with the tidy verse

```{r}
EC50.data %>%
  group_by(is) %>%
  nest() %>%
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc, 
                              fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                         names = c("Slope", "Lower", "Upper", "EC50"))))) %>%
  mutate(ec50 = map(ll.4.mod, ~ED(., 
                              respLev = c(50), 
                              type = "relative",
                              interval = "delta")[[1]])) %>%
  unnest(ec50)
```



