---
title: "Data Wrangling"
author: "Sinka Khadijah Abubakar"
date: "2026-03-08"
output: 
  html_document:
    toc: true
    toc_float: true
  pdf_document:
  md_document:
    variant: gfm
  word_document:
---


Question 1. Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read them into R using relative file paths.

```{r}
# loading reuired packages
library(tidyverse)

# Reading csv files
diversity <- read.csv("DiversityData.csv")
metadata <- read.csv("Metadata.csv")
```

Question 2. Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe alpha.

```{r}
# Joining the two dataframes together by the common column ‘Code’
alpha <- left_join(diversity, metadata, by = "Code")
head(alpha)
```

Question 3. Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column.
a. Using mutate, create a new column to calculate Pielou’s evenness index.
b. Name the resulting dataframe alpha_even.

```{r}
# Creating a new dataframe and calculating Pielou’s evenness index
alpha_even <- alpha %>%
  mutate(even = shannon / log(richness))
head(alpha_even)
```


Question 4. Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
a. Start with the alpha_even dataframe
b. Group the data: group the data by Crop and Time_Point.
c. Summarize the data: Calculate the mean, count, standard deviation, and standard error
for the even variable within each group.
d. Name the resulting dataframe alpha_average

```{r}
alpha_average <- alpha_even %>%
  group_by(Crop, Time_Point) %>%    # grouping the data by Crop and Time_Point
  summarise(                        # Summarizing the data
    mean.even = mean(even),
    n = n(),
    sd = sd(even),
    se = sd / sqrt(n)
  )
head(alpha_average)              # plotting the head of the resulting dataframe
```


Question 5. Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
a. Start with the alpha_average dataframe
b. Select relevant columns: select the columns Time_Point, Crop, and mean.even.
c. Reshape the data: Use the pivot_wider function to transform the data from long to wide
format, creating new columns for each Crop with values from mean.even.
d. Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even
by calculating the difference between Soil and Cotton, and Soil and Soybean,
respectively.
e. Name the resulting dataframe alpha_average2

```{r}
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, mean.even) %>%   # selecting relevant columns
  pivot_wider(names_from = Crop, values_from = mean.even) %>%   # pivoting wider
  mutate(                                   # calculating differences
    diff.cotton.even = Soil - Cotton,
    diff.soybean.even = Soil - Soybean
  )
head(alpha_average2)                # plotting the head of the resulting dataframe
```


Question6. Connecting it to plots
a. Start with the alpha_average2 dataframe
b. Select relevant columns: select the columns Time_Point, diff.cotton.even, and
diff.soybean.even.
c. Reshape the data: Use the pivot_longer function to transform the data from wide to long
format, creating a new column named diff that contains the values from diff.cotton.even
and diff.soybean.even.
d. Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column
‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named
‘values’ come from the pivot_longer. The resulting plot should look like the one to the
right.

```{r}
plot_data <- alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%   # selecting columns
  pivot_longer(                                 # pivoting longer
    c(diff.cotton.even, diff.soybean.even),     # creating new column diff
    names_to = "diff"
  )

library(ggplot2)                         # plotting
ggplot(plot_data, aes(x = Time_Point, y = value, color = diff)) +
  geom_line(size = 1.2) +
  xlab("Time (hrs)") +
  ylab("Difference from Soil in Pielou's Evenness") +
  theme_minimal()
```


Question 7. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 5. Provide me a link to your github written as a clickable link in your .pdf or .docx

**GitHub Repository**

[View my Coding Challenge 5 on GitHub](https://github.com/sza0209/PLAP/blob/main/Assignments/Coding%20Challenge5/Data-Wrangling.md)
