---
title: "Linear Model"
author: "Sinka Khadijah Abubakar"
date: "2026-03-26"
output: 
  html_document:
  pdf_document:
  md_document:
    variant: gfm
  word_document:
---

Question 1 (4 pts).
Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting and Rep into factors using the function as.factor

```{r}
# Loading libraries
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
```


```{r}
# reading in the data
STAND <- read.csv("PlantEmergence.csv")
```


```{r}
# Converting variables into factors
STAND$Treatment <- as.factor(STAND$Treatment)
STAND$DaysAfterPlanting <- as.factor(STAND$DaysAfterPlanting)
STAND$Rep <- as.factor(STAND$Rep)
```


Question 2 (5 pts)
Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results.

```{r}
# fitting a linear model with an interaction term
lm.emerge <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = STAND)
```


```{r}
# providing the summary
summary(lm.emerge)
```


```{r}
# running ANOVA
anova(lm.emerge)
```


Question 3 (5 pts)
Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

Answer
The interaction between Treatment and DaysAfterPlanting was not significant (ANOVA, p = 1.00), indicating that the effect of treatment did not change over time. And since the interaction is not significant (p = 1), we do not need to include the interaction term in the final model.


```{r}
# Simplified linear model without the interaction term
lm.simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = STAND)

# summary 
summary(lm.simple)

#anova
anova(lm.simple)
```

Interpretation of the intercept and the coefficient for Treatment 2

The coefficient for Treatment 2 (-134.53) indicates that emergence in Treatment 2 was approximately 135 plants lower than in Treatment(p < 2e-16).


Question 4 (5 pts)
Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function.
Interpret the results.

```{r}
# Calculating least square means for Treatment
lsm.trt <- emmeans(lm.simple, ~ Treatment)
lsm.trt
```

```{r}
# performing a Tukey separation with the compact letter display
library(multcompView)
tukey.trt <- cld(lsm.trt, alpha = 0.05, Letters = letters, adjust = "tukey")
tukey.trt
```

Interpretation
Least square means were calculated to estimate the average plant emergence for each treatment while accounting for variation across days after planting. The adjusted means showed that Treatment 2 had substantially lower emergence (mean ≈ 56 plants) compared to all other treatments, which ranged from approximately 188 to 201 plants.

Tukey’s multiple comparison test with compact letter display confirmed these differences. Treatment 2 was assigned to its own group (“a”), indicating that it was significantly different from all other treatments. Treatments 3, 5, and 6 formed the highest-performing group (“c”), while Treatments 1, 4, 8, and 9 were intermediate (“bc”) and did not differ significantly from either the high or moderate groups. Treatment 7 formed a slightly lower group (“b”) but was still significantly higher than Treatment 2.



Question 5 (4 pts)
The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters.

```{r}
# function 
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]
  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details =
TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1],
                                str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) +
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") +
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
return(plot)
}
```


```{r}
# Generate the plot using simplified model and Treatment as the factor
plot_cldbars_onefactor(lm.simple, "Treatment")
```


Significance of the letters
Letters above each bar represent results of Tukey-adjusted pairwise comparisons of the least square means. Treatments that share the same letter are NOT significantly different from each other.Treatments with different letters are significantly different. 



Question 6 (2 pts)
Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

**GitHub Repository**

[View my Coding Challenge 7 on GitHub](https://github.com/sza0209/PLAP/blob/main/Assignments/LinearModels/LinearModelCodingChallenge.md)
