# Data Visualization 1 Assignment

# Name: Sinka Khadijah Abubakar

### Question 1 (5 pts.)

Explain the following concepts about ggplot

a. What three elements do you need to produce a ggplot?
Answer: (1) a dataset 
        (2) aesthetic mappings defined with aes() (such as x and y variables)
        (3) at least one geom to tell ggplot how to draw the data.

b. What is a geom?
Answer: A geom is the geometric object that represents the data visually, such as points (geom_point()), boxes (geom_boxplot()), bars (geom_bar() or stat_summary()), or lines (geom_line()).

c. What is a facet?
Answer: A facet splits the data into multiple panels based on the levels of a categorical variable, allowing comparisons across groups using the same axes.

d. Explain the concept of layering.
Answer: Layering means building a plot by adding multiple components (geoms, scales, labels) using the "+" sign, with each layer drawn in order from background to foreground.

e. Where do you add x and y variables and map different shapes, colors, and other attributes to the data?
Answer: X and y variables and aesthetic mappings are added inside the aes() function, either in the main ggplot() call or within individual geoms.



### Question 2 (4 pts.)

Make a boxplot using ggplot with DON as the y variable, treatment as the x
variable, and color mapped to the wheat cultivar. Change the y label to “DON (ppm)” and
make the x label blank.

```{r}
# Loading ggplot2 library
library(ggplot2)

# Loading and checking mycotoxin data
mycotoxin <- read.csv("MycotoxinData.csv")
str(mycotoxin)
```

```{r}
# Changing DON from character to numeric vaiable
mycotoxin$DON = as.numeric(mycotoxin$DON)

# Making a boxplot using ggplot with DON as the y variable, treatment as the x variable, and color mapped to the wheat cultivar
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar)) +
  geom_boxplot() +
  xlab("") +              # leaving the x label blank
  ylab("DON (ppm)")       # Changing the y label to “DON (ppm)”
```


### Question 3 (4 pts.)

Now convert this data into a bar chart with standard-error error bars using the stat_summary() command. (hint: use position = dodge)

```{r}
# Plotting the data as a bar chart with standard error bars
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  xlab("") +
  ylab("DON (ppm)")
```



### Question 4 (3 pts.)

Add points to the foreground of the boxplot and bar chart you made in questions 2 & 3 that show the distribution of points. Set the shape = 21 and the outline color black (hint: use jitter_dodge).


```{r}
# Box plot with points to the foreground
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(dodge.width = 0.9),shape = 21,color = "black") +  ## adding points to the foreground, setting the shape = 21 and the outline color black
  xlab("") +
  ylab("DON (ppm)")
```


```{r}
# Bar chart with points to the foreground
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  geom_point(position = position_jitterdodge(dodge.width = 0.9),shape = 21,color = "black") +  ## adding points to the foreground, setting the shape = 21 and the outline color black
  xlab("") +
  ylab("DON (ppm)")
```


### Question 5 (2 pts.)
Add a facet to the plots based on cultivar.

```{r}
# Adding facet to box plot based on cultivar
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(dodge.width = 0.9),shape = 21,color = "black") +
  facet_wrap(~Cultivar) +    ## adding facet based on cultivar
  xlab("") +
  ylab("DON (ppm)")
```


```{r}
# Adding facet to box plot based on cultivar
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  geom_point(position = position_jitterdodge(dodge.width = 0.9),shape = 21,color = "black") +
   facet_wrap(~Cultivar) +     ## adding facet based on cultivar
  xlab("") +
  ylab("DON (ppm)")
```


### Question 6 (3 pts.)

Add transparency to the points so you can still see the boxplot or bar in the
background.

```{r}
# Making points on the boxplot transparent to see the boxplot
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(dodge.width = 0.9),shape = 21,color = "black", alpha = 0.4) +  # adding "alpha = 0.4" argument to make points transparent
  facet_wrap(~Cultivar) +
  xlab("") +
  ylab("DON (ppm)")
```



```{r}
# Making points on the bar chart transparent to see the bar chart
ggplot(mycotoxin, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  geom_point(position = position_jitterdodge(dodge.width = 0.9),shape = 21,color = "black", alpha = 0.4) +  # adding "alpha = 0.4" argument to make points transparent
   facet_wrap(~Cultivar) +
  xlab("") +
  ylab("DON (ppm)")
```

