# Intro to R Assignment

# Name: Sinka Khadijah Abubakar

### Question 1 (5 pts.)
Explain the following things about R and Rstudio: 

o	Where do you type code? 
Answer: Script Editor (usually top-left pane)

o	Where is the output of the code? 
Answer: Console (usually bottom left pane)

o	Where do plots show up?
Answer: In the Plots pane (usually bottom-right).

o	What is a global environment? 
Answer: This is where R stores objects (data, variables, functions) that we create.

o	How do you get help for a function through R studio? 
Answer: Use ?functionName or help(functionName).

o	What is an R package? 
Answer: It is simply a collection of functions, data, and documentation

o	What is a function? 
Answer: It is simply a reusable block of code that performs a specific task.

o	How do you find the installed and loaded packages? 
Answer: In the Packages pane (usually bottom-right, next to plots). All installed packages would show up in there and the loaded ones would be checked.

o	What is a working directory, and how do you find it? 
Answer: This is the folder that R uses to read and save files, and yoy can find it with getwd().

o	What is a relative file path and how is it different than an absolute file path? 
Answer: A relative path is based on the working directory, while an absolute path gives the full location from the root folder of your computer.


### Question 2 (2 pts.)
Explain the steps to start a new R project connected to a GitHub repository

Create a new repository on GitHub.Open RStudio and create a new project. To do this, in RStudio, go to File > New Project > Version Control > Git.Copy and paste the repository URL and choose a folder on your computer to save the project.  


### Question 3 (3 pts.)
Please explain the different data classes and how they are similar and different from each other.

o	Vector
Answer: A 1-dimensional collection of elements of the same type (numeric, character, logical, etc.).

o	Dataframe
Answer: A 2-dimensional table where columns can be different types (numeric, character, logical, etc.).

o	Matrix
Answer: A 2-dimensional array (rows and columns) where all elements must be the same type.

All three store multiple values and can be indexed, but vectors are 1D, while matrices and data frames are 2D.


### Question 4 (10 pts.)
Complete all tasks in an R script and push it to GitHub.

o	Create a vector named 'z' with the values 1 to 200

```{r}
z = 1:200
```

o	Print the mean and standard deviation of z on the console

```{r}
mean(z)
sd(z)
```

o	Create a logical vector named zlog that is 'TRUE' for z values greater than 30 and 'FALSE' otherwise.

```{r}
zlog = z > 30
```

o	Make a dataframe with z and zlog as columns. Name the dataframe zdf

```{r}
zdf = data.frame(z, zlog)
```

o	Change the column names in your new dataframe to equal “zvec” and “zlogic”

```{r}
 colnames(zdf) = c("zvec", "zlogic")
```

o	Make a new column in your dataframe equal to zvec squared (i.e., z2). Call the new column zsquared. 

```{r}
zdf$zsquared <- zdf$zvec^2
```

o	Subset the dataframe with and without the subset() function to only include values of zsquared greater than 10 and less than 100 

```{r}
subset(zdf, zsquared > 10 & zsquared < 100)
```


```{r}
zdf[zdf$zsquared > 10 & zdf$zsquared < 100, ]
```


o	Subset the zdf dataframe to only include the values on row 26

```{r}
zdf[26, ]
```

o	Subset the zdf dataframe to only include the values in the column zsquared in the 180th row.

```{r}
zdf$zsquared[180]
```


